---
title: "bank customer churn"
output: 
  html_document: 
    keep_md: yes
---

```{r message = FALSE, warning=FALSE}

# install.packages("tidyverse")
# install.packages("ggpubr")
# install.packages("ggcorrplot")
library(tidyverse)
# tidyverse_update()
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(dplyr)
```




```{r message = FALSE, warning=FALSE}
setwd("D:/Git/Project/Bank_churn/")
```

```{r}
df<-read.csv("Bank_Churn_Modelling.csv")
str(df)
summary(df)
print("null values: ")
colSums(is.na(df)) 
```

```{r}

df <- df %>%
  dplyr::select(-c(RowNumber,CustomerId,Surname)) %>%
  mutate(Geography=as.factor(Geography),
         Gender=as.factor(Gender),
         Tenure=as.factor(Tenure),
         NumOfProducts=as.factor(NumOfProducts),
         HasCrCard = as.factor(HasCrCard),
         IsActiveMember = as.factor(IsActiveMember),
         Exited=as.factor(Exited))
```

x<-is.null(df)
print(paste("null values: ",is.null(df)))


bank %>% 
select_if(function(col) is.numeric(col) | all(col == .$Churn)) %>%
gather(key = "Variable", value = "Value", -Churn) %>% ggplot(aes(Value, fill = Churn)) + 
geom_histogram(bins = 20) + 
facet_wrap(~ Variable, scales = "free") + 
labs( title = "Numerical variables histograms", x = "" ) + scale_fill_economist()

```{r}
df %>%
  select(where(is.numeric),Exited)%>%
  gather(key = "Variable", value = "Value",-Exited) %>%
  ggplot(aes(Value,fill=Exited))+
    facet_wrap(~Variable,scales="free")+
    geom_histogram(bins=30,color='black')+
    scale_fill_manual(values=c("cornflowerblue","brown"))
df%>%
  keep(is.numeric)%>%summary()


```


```{r}
Age_box<-ggplot(data=df)+
  geom_boxplot(mapping = aes(Exited,Age,fill=Exited),width=.5)+
  scale_fill_manual(values=c("cornflowerblue","brown"))


Balance_box<-ggplot(data=df)+
  geom_boxplot(mapping = aes(Exited,Balance,fill=Exited),width=.5)+
  scale_fill_manual(values=c("cornflowerblue","brown"))

CreditScore_box<-ggplot(data=df)+
  geom_boxplot(mapping = aes(Exited,CreditScore,fill=Exited),width=.5)+
  scale_fill_manual(values=c("cornflowerblue","brown"))
                    
                    
EstimatedSalary_box<-ggplot(data=df)+
  geom_boxplot(mapping = aes(Exited,EstimatedSalary,fill=Exited),width=.5)+
  scale_fill_manual(values=c("cornflowerblue","brown"))

ggarrange(Age_box,Balance_box, CreditScore_box,EstimatedSalary_box,
         ncol = 2, nrow = 2,common.legend = TRUE)

```

## Categorical 

**Existed (Churn)** defined as predicted variable

```{r}
ggplot(df,aes(Exited))+
  geom_bar(fill=c("cornflowerblue","brown"),width=0.4)+
  coord_flip()+
  theme(aspect.ratio = 0.3)

    

```

**AGE**
```{r}
# df$Age_cat<-cut(df$Age, breaks=c(25,35,45,55,65,75,85), right = FALSE)
df$Age_cat<-cut(df$Age, breaks=c(0,20,40,60,80,100), right = FALSE)
df$Age_cat<-as.factor(df$Age_cat)
# df$Age_cat<-factor(df$Age_cat,levels=c(1,2,3,4,5),labels=c("0-19","20-39","40-59","60-79","80-99"))
# 
glimpse(df)

```


```{r}
Agecat_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(Age_cat,fill=Exited),position='fill',width=0.5)+
  scale_x_discrete("Age_cat")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
  coord_flip()+
  theme(aspect.ratio = 1)
Agecat_cat<-ggplot(df,aes(Age_cat))+
  geom_bar(fill="grey41",width=0.5)+
  coord_flip()+
  theme(aspect.ratio = 0.7)

ggarrange(Agecat_churn, Agecat_cat,
         ncol = 2, nrow = 1)
```


**Gender**

```{r}

Gender_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(Gender,fill=Exited),position='fill',width=0.5)+
  scale_x_discrete("Gender")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
  coord_flip()+
  theme(aspect.ratio = 0.7)
Gender_cat<-ggplot(df,aes(Gender))+
  geom_bar(fill="grey41",width=0.5)+
  coord_flip()+
  theme(aspect.ratio = 0.5)
# Gender_cat2<-ggplot(data=df)+
#   geom_bar(mapping = aes(Gender,fill=Exited),width=0.5)+
#   scale_x_discrete("Gender")+
#   scale_y_continuous("Percent")+
#   guides(fill=guide_legend(title="Churn"))+
#   scale_fill_manual(values=c("grey41","brown"))+
#   coord_flip()+
#   theme(aspect.ratio = 0.7)

ggarrange(Gender_churn, Gender_cat,
         ncol = 2, nrow = 1)


                   
```


**Location**
```{r}
Location_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(Geography,fill=Exited),position='fill',width=0.5)+
  scale_x_discrete("Geography")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
  coord_flip()+
  theme(aspect.ratio = 1)
Location_cat<-ggplot(df,aes(Geography))+
  geom_bar(fill="grey41",width=0.5)+
  coord_flip()+
  theme(aspect.ratio = 0.7)

ggarrange(Location_churn, Location_cat,
         ncol = 2, nrow = 1)
```


**Number of Products**
```{r}

Product_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(factor(NumOfProducts),fill=factor(Exited)),position='fill',width=0.7)+
  scale_x_discrete("NumofProducts")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
  coord_flip()+
  theme(aspect.ratio = 0.8)
Product_cat<-ggplot(df,aes(NumOfProducts))+
  geom_bar(fill="grey41",width=0.6)+
  coord_flip()+
  theme(aspect.ratio = 0.6)

ggarrange(Product_churn, Product_cat,
         ncol = 2, nrow = 1)
  
```

**Credit Card**
```{r}
Creditcard_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(factor(HasCrCard),fill=factor(Exited)),position='fill',width=0.4)+
  scale_x_discrete("CreditCard")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
  coord_flip()+
  theme(aspect.ratio = 1)
  
Creditcard_cat<-ggplot(df,aes(factor(HasCrCard)))+
  geom_bar(width=0.4,fill='grey41')+
  theme(legend.position = "left")+
  coord_flip()+
  theme(aspect.ratio = 0.8)

ggarrange(Creditcard_churn, Creditcard_cat,
         ncol = 2, nrow = 1)    

```
**Active Member**
```{r}

Active_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(IsActiveMember,fill=Exited),position='fill',width=0.4)+
  scale_x_discrete("IsActiveMember")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  coord_flip()+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
 theme(aspect.ratio = 1)

Active_cat<-ggplot(df,aes(IsActiveMember))+
  geom_bar(width=0.4,fill='grey51')+
  theme(legend.position = "left")+
  coord_flip()+
  theme(aspect.ratio = 0.8)
ggarrange(Active_churn, Active_cat,
         ncol = 2, nrow = 1)
```

**Tenure**
```{r}

Tenure_churn<-ggplot(data=df)+
  geom_bar(mapping = aes(Tenure,fill=Exited),position='fill',width=0.5)+
  scale_x_discrete("Tenure")+
  scale_y_continuous("Percent")+
  guides(fill=guide_legend(title="Churn"))+
  scale_fill_manual(values=c("cornflowerblue","brown"))+
 theme(aspect.ratio = 1)

Tenure_cat<-ggplot(df,aes(Tenure))+
  geom_bar(width=0.5,fill='grey51')+
  theme(legend.position = "left")+
  theme(aspect.ratio = 0.8)
ggarrange(Tenure_churn,Tenure_cat,
         ncol = 2, nrow = 1)
```
## chi-square test of independence

**Gender**
```{r}
gender_table<- table(df$Gender,df$Exited,dnn=c('Gender','Churn'))
round(prop.table(gender_table,margin =2),4)*100 #col margin
round(prop.table(gender_table,margin =1),4)*100 #row margin
chisq.test(df$Gender,df$Exited,correct=FALSE)
```

**Credit Card**

```{r}

creditcard_table<- table(df$HasCrCard,df$Exited,dnn=c('CreditCard','Churn'))
round(prop.table(creditcard_table,margin =2),4)*100 #col margin
round(prop.table(creditcard_table,margin =1),4)*100 #row margin
chisq.test(df$HasCrCard,df$Exited,correct=FALSE)
```
**Num of Products**

```{r}
table(df$NumOfProducts,df$Exited)
product_table<- table(df$NumOfProducts,df$Exited,dnn=c('#Product','Churn'))
round(prop.table(product_table,margin =2),4)*100 #col margin
round(prop.table(product_table,margin =1),4)*100 #row margin
chisq.test(df$NumOfProducts,df$Exited,correct=FALSE)
```

**Active Member**

```{r}
table(df$IsActiveMember,df$Exited)
active_table<- table(df$IsActiveMember,df$Exited,dnn=c('Active','Churn'))
round(prop.table(active_table,margin =2),4)*100 #col margin
round(prop.table(active_table,margin =1),4)*100 #row margin
chisq.test(df$IsActiveMember,df$Exited,correct=FALSE)
```

**Location**
```{r}
table(df$Geography,df$Exited)

location_table<- table(df$Geography,df$Exited,dnn=c('Location','Churn'))
round(prop.table(location_table,margin =2),4)*100 #col margin
round(prop.table(location_table,margin =1),4)*100 #row margin
chisq.test(df$Geography,df$Exited,correct=FALSE)
```

**Tenure**
```{r}
table(df$Tenure,df$Exited)
tenure_table<- table(df$Tenure,df$Exited,dnn=c('Tenure','Churn'))
round(prop.table(tenure_table,margin =2),4)*100 #col margin
round(prop.table(tenure_table,margin =1),4)*100 #row margin
chisq.test(df$Tenure,df$Exited,correct=FALSE)
```
-c(HasCrCard,Tenure)


```{r}
df_model <- df %>%
  dplyr::select(-c(HasCrCard,Tenure)) 

```


```{r}
df %>%
  select(where(is.numeric)) %>%
  pairs()
```

**Logistic regression**
```{r}
logis_mod01<- glm(Exited ~., data = df, family = binomial)
summary(logis_mod01)
```
```{r}
logis_mod02<- glm(Exited ~.-Tenure-HasCrCard-CreditScore, data =df, family = binomial)
summary(logis_mod02)

```

```{r}
anova(logis_mod02,test='Chisq')
```
```{r}
logis_mod03<- glm(Exited ~Age+NumOfProducts+Gender+Geography+IsActiveMember, data =df, family = binomial)
summary(logis_mod03)

```
```{r}
anova(logis_mod03,test='Chisq')
`````{r}
lm_01 <- glm(Exited ~., data = df_model, family = binomial)
summary(lm_01)
````






```{r}
library(caTools)
set.seed(123)
sample_data = sample.split(df_model, SplitRatio = 0.8)
train_data <- subset(df_model, sample_data == TRUE)
test_data <- subset(df_model, sample_data == FALSE)

```

```{r}
#Logistic regression(full model)
logis_full<- glm(Exited ~., data = train_data, family = binomial)
summary(logis_full)
anova(logis_full,test='Chisq')
```

```{r}
#result_#Logistic regression(full model)
result_full<-predict(logis_full,test_data,type="response")
predict_full<-ifelse(result_full>0.5,"1","0")
(confusion<-table(predict=predict_full,actual=test_data$Exited))
full_accuracy<-round(mean(predict_full==test_data$Exited),3) #model accuracy
print(paste('Logistic Regression (full) Accuracy',full_accuracy))
```


```{r}
library(MASS)
logis_step <-logis_full %>% stepAIC(trace = FALSE)
coef(logis_step)
```

```{r}
#predictions
result_step <- predict(logis_step,test_data,type="response")
predict_step <- ifelse(result_step >0.5,"1","0")
table(Predict=predict_step,Actual=test_data$Exited)
step_accuracy<-round(mean(predict_step==test_data$Exited),3)#accuracy
print(paste('Logistic Regression (Stepwise) Accuracy',step_accuracy))
      
```


```{r}
library(caret)
train_ctr <- trainControl(method = 'cv', number = 10, classProbs = TRUE, summaryFunction = twoClassSummary)
Logistic_model_t <- train(Exited ~ .,data = train_data, method = 'glm',family = 'binomial',trControl = train_ctr, metric = 'ROC')
```

```{r}
# install.packages("randomForest")
library(randomForest)
```

```{r}
# load caret
library(caret)# define parameters used in training data
my_control <- trainControl(method = "repeatedcv", 
                           number = 5,
                           repeats = 5, 
                           verboseIter = TRUE, 
                           summaryFunction = twoClassSummary, 
                           classProbs = TRUE)
# train random forest

```



```{r}
rf_full<- randomForest(Exited ~., data = train_data)
print(rf_full)
```




```{r}
# install.packages('rpart')
# install.packages('partykit')
# install.packages('caTools')
library(caTools)
library(rpart)
library(rpart.plot)
library(party)
library(partykit)
```



Decision Tree
```{r}
set.seed(123)
sample_data = sample.split(df_model, SplitRatio = 0.75)
train_data <- subset(df_model, sample_data == TRUE)
test_data <- subset(df_model, sample_data == FALSE)

rtree <- rpart(Exited ~ ., data=train_data)
rpart.plot(rtree)
```
```{r}
ctree_ <- ctree(Exited ~ ., train_data)
plot(ctree_)
```


```{r}
# Predict the values of the test set: pred
pred <- predict(rtree, test_data, type = "class")

# Construct the confusion matrix: conf
confusion <- table(test_data$Exited, pred)

# Print out the accuracy
print(confusion)
print(sum(diag(confusion)) / sum(confusion))

```
random forest

```{r}
# install.packages("randomForest")
library(randomForest)
```

```{r}

rfmod <- randomForest(Exited ~ ., data = train_data, importance = T)
rfmod
predTrain <- predict(rfmod, train_data, type = "class")
table(predTrain, train_data$Exited)
predTest <- predict(rfmod, test_data, type = "class")
rfacc <- mean(predTest == test_data$Exited)
rfacc
table(predTest, test_data$Exited)

# Construct the confusion matrix: conf
confusion <- table(test_data$Exited, predTest)

# Print out the accuracy
print(confusion)
print(sum(diag(confusion)) / sum(confusion))
```

